### 2012-05-09 GUIErrorDetector/paper/introduction.tex
Mike's edits through section 4.1

End-user satisfaction depends in part on the responsiveness and
robustness of a software application's GUI.

@@

%\subsection{The Single-thread Rule for GUI Object Access}
%ugly hack
%\subsection{\hspace{-1.1ex}\mbox{The}~\mbox{Single-thread}~\mbox{Rule}~\mbox{for}~\mbox{GUI}~\mbox{Object}~\mbox{Access}}

\subsection{The Single-thread Rule}

@@

The single-GUI-thread rule requires GUI application developers to
 ensure that all GUI objects are accessed only by the UI thread.
If not, an \textit{invalid thread access error} will occur. This may
terminate the application, but it is considered preferable
to nondeterministic concurrency errors.

@@

Take the popular Standard Widget Toolkit (SWT) GUI framework as an example. 
Invalid thread access is one of the top 3 bugs in developing a SWT application, and is
the source of many concurrency bugs~\cite{top3bugs}.
A Google search for ``SWTException:Invalid thread access''  returns over 11,800 entries,
consisting of numerous bug reports, forum posts, and mailing list threads on
this problem. Eclipse, the IDE for Java development, is built on top of SWT\@.
Searching for ``SWTException:Invalid thread access'' in Eclipse's bug repository
and discussion forum returns over 2732 bug reports and 351 distinct discussion threads, respectively. 
We manually studied all 156 distinct \textit{confirmed} bug reports, and 
found this error has been confirmed in at least 20 distinct Eclipse projects
and 40 distinct Eclipse components. Even after over 10 years of active development,
a recent release of Eclipse still contains this error (bug id: 333533, reported in January 2011).
 In addition, the invalid thread access error
is severe. It is user-perceivable, it cannot be recovered by the program itself,
and it often terminates the whole application. In many circumstances as described in the bug reports,
users must restart the application to recover from the error.
Furthermore, many reported  errors are non-trivial to diagnose and fix.
Developers usually need non-local reasoning to find the specific
UI interactions that can trigger the bug;  it took developers 2 years
to fix Eclipse bug 51757 and verify the patch.

@@

%Also, you've got no confidence that your app will continue to work on future CPUs with more and more cores - which are more prone to encountering weird threading issues due to them being truely concurrent rather than just simulated by the OS.

@@

%Software testing, the most widely used method to ensure program correctness,
%is insufficient to detect such errors. 
It is infeasible for testing to cover the enormous space of possible interactions
with a GUI. Each sequence of GUI events can result in
a different state, and each GUI event may need to be evaluated in all of
these states. A  software system like Eclipse
often has a test suite that achieves fairly high statement coverage,
but many paths executed by bug-triggering UI sequences are still not covered.

Stylized coding patterns are also inadequate. One possible rule is to always
access GUI objects via asynchronous message passing, to ensure a GUI object is accessed in
the UI thread. For example, a developer could have prevented the bug in Figure~\ref{fig:androiderror}
by wrapping line 222 inside a \CodeIn{post} message-passing method\footnote{The
\CodeIn{post} method in class \CodeIn{android.widget.View} is
a standard way to send asynchronous messages to the UI-thread.}, as follows:

@@

Such an approach is desirable for accesses from non-UI threads, but it is unnecessary.
In our evaluation on real-world programs, we found that a simple analysis that
requires GUI operations to be in a message issued
an unacceptable number of warnings. %false positives and redundant warnings.
Furthermore, this approach is dangerous for accesses from the UI thread. Asynchronized message passing offers
no timing guarantee, so a GUI object may have already been
disposed before the message sent to it arrives, causing other bugs.

@@

Static analysis may report false positives due to its
conservative nature, or multiple warnings that actually correspond
to the same error. To address such limitations, we devise a set
 of error filters to remove likely false positives and redundant warnings.
The filters introduce potential unsoundness to our algorithm, but in practice
they work well and make our technique more usable.

@@

We implemented an invalid-thread-access-error detection tool that supports
four popular GUI frameworks: SWT, Eclipse plugin framework,
Swing, and Android. Swing and SWT are the
two dominant GUI frameworks for desktop Java applications.
Eclipse is a widely-used IDE for Java. Android is the \#1 platform for
mobile applications with market share 56\% as of September 2011.
Although our technique is applicable to any GUI
framework with the single-GUI-thread rule, each framework
has its own definition of \textit{UI thread} and \textit{program 
entry points}. Thus, our implementation is parameterized with respect to
those framework-specific parts (see Section~\ref{sec:platforms}).

Our tool works in an automatic manner,
and scales to realistic programs.
We evaluated our implementation on \subnum programs comprising \totaloc LOC.
The experimental results demonstrate
that: our technique is effective (it found \bugs real-world errors
 and produced only \falses false positive warnings);
our reflection-aware call graph construction algorithm helps in
finding errors in Android applications;
and our proposed error filters significantly reduce the number of warnings.
%false positives and redundant warnings.

@@

\item \textbf{Technique.} We formulate finding
the invalid thread access error as a call graph reachability problem,
and present a general error detection technique.
In addition, we use a reflection-aware
call graph construction algorithm (Section~\ref{sec:technique}).

@@

\item \textbf{Evaluation.} We applied our tool to \subnum programs 
from 4 different frameworks, comprising \totaloc LOC. The results
show the usefulness of the proposed technique (Section~\ref{sec:evaluation}).

@@

>>>>

End-user satisfaction depends in part on the responsiveness and
robustness of a software application's GUI\@.

@@

%\subsection{The Single-GUI-Thread Rule for GUI Object Access}
%ugly hack
%\subsection{\hspace{-1.1ex}\mbox{The}~\mbox{Single-thread}~\mbox{Rule}~\mbox{for}~\mbox{GUI}~\mbox{Object}~\mbox{Access}}

\subsection{The Single-GUI-Thread Rule}

@@

The single-GUI-thread rule requires GUI application developers to
 ensure that all GUI objects are accessed only by the UI thread.
If not, an \textit{invalid thread access error} will occur. This may
terminate the application --- doing so is considered preferable
to nondeterministic concurrency errors.

@@

Take the popular Standard Widget Toolkit (SWT) GUI framework as an example. 
Invalid thread access is one of the top 3 bugs in developing a SWT application, and is
the source of many concurrency bugs~\cite{top3bugs}.
A Google search for ``SWTException:Invalid thread access''  returns over 11,800 hits,
consisting of numerous bug reports, forum posts, and mailing list threads on
this problem. Eclipse, the IDE for Java development, is built on top of SWT\@.
Searching for ``SWTException:Invalid thread access'' in Eclipse's bug repository
and discussion forum returns over 2732 bug reports and 351 distinct discussion threads, respectively. 
We manually studied all 156 distinct \textit{confirmed} bug reports, and 
found this error has been confirmed in at least 20 distinct Eclipse projects
and 40 distinct Eclipse components. Even after over 10 years of active development,
a recent release of Eclipse still contains this error (bug id: 333533, reported in January 2011).
 In addition, the invalid thread access error
is severe. It is user-perceivable, it cannot be recovered by the program itself,
and it often terminates the whole application. In many circumstances as described in the bug reports,
users must restart the application to recover from the error.
Furthermore, many reported  errors are non-trivial to diagnose and fix.
Developers usually need non-local reasoning to find the specific
UI interactions that can trigger the bug;  it took developers 2 years
to fix Eclipse bug 51757 and verify the patch.

@@

%Also, you've got no confidence that your app will continue to work on future CPUs with more and more cores - which are more prone to encountering weird threading issues due to them being truely concurrent rather than just simulated by the OS\@.

@@

%Software testing, the most widely used method to ensure program correctness,
%is insufficient to detect such errors. 
It is infeasible for testing to cover the enormous space of possible interactions
with a GUI\@. Each sequence of GUI events can result in
a different state, and each GUI event may need to be evaluated in all of
these states. A  software system like Eclipse
often has a test suite that achieves fairly high statement coverage,
but many paths executed by bug-triggering UI sequences are still not covered.

Stylized coding patterns are also inadequate. One possible rule is to always
access GUI objects via asynchronous message passing, to ensure a GUI object is accessed in
the UI thread. For example, a developer could have prevented the bug in Figure~\ref{fig:androiderror}
by wrapping line 222 inside a \CodeIn{post} message-passing method\footnote{The
\CodeIn{post} method in class \CodeIn{android.widget.View} is
a standard way to send asynchronous messages to the UI thread.}, as follows:

@@

Such an approach is desirable for accesses from non-UI threads, but it is
not necessary for all accesses.
In our evaluation on real-world programs, we found that a simple analysis that
requires GUI operations to be in a message issued
an unacceptable number of warnings. %false positives and redundant warnings.
Furthermore, this approach is dangerous for accesses from the UI thread. Asynchronous message passing offers
no timing guarantee, so a GUI object may have already been
disposed before the message sent to it arrives, causing other bugs.

@@

Static analysis may report false positives due to its
conservative nature, or may report multiple warnings that actually correspond
to the same error. To address such limitations, we devised a set
 of error filters to remove likely false positives and redundant warnings.
The filters introduce potential unsoundness to our algorithm, but in practice
they work well and make our technique more usable.

@@

We implemented an invalid-thread-access-error detection tool that supports
four popular GUI frameworks: SWT, Eclipse plugin framework,
Swing, and Android. Swing and SWT are the
two dominant GUI frameworks for desktop Java applications.
Eclipse is the most widely-used IDE for Java. Android is the \#1 platform for
mobile applications with market share 56\% as of September 2011.
Although our technique is applicable to any GUI
framework with the single-GUI-thread rule, each framework
has its own definition of \textit{UI thread} and \textit{program 
entry points}. Thus, our implementation is parameterized with respect to
those framework-specific parts (see Section~\ref{sec:platforms}).

Our tool works in an automatic manner
and scales to realistic programs.
We evaluated our implementation on \subnum programs comprising \totaloc LOC\@.
The experimental results demonstrate
that: our technique is effective (it found \bugs real-world errors
 and produced only \falses false positive warnings);
our reflection-aware call graph construction algorithm helps in
finding errors in Android applications;
and our proposed error filters significantly reduce the number of warnings
and thus the programmer effort.
%false positives and redundant warnings.

@@

\item \textbf{Technique.} We formulate finding
invalid thread access errors as an elegant call graph reachability problem,
and present a general error detection technique.
In addition, we use a reflection-aware
call graph construction algorithm (Section~\ref{sec:technique}).

@@

\item \textbf{Evaluation.} We applied our tool to \subnum programs 
from 4 different frameworks, comprising \totaloc LOC\@. The results
show the usefulness of the proposed technique (Section~\ref{sec:evaluation}).

@@

%  LocalWords:  multithreaded SWT plugin MacOS myquotation 0pt 10pt GL10
%  LocalWords:  SWTException LayerView LayerController mRenderer mGLThread
%  LocalWords:  LayerRenderer setRenderer GLThread onSurfaceChanged gl RTA
%  LocalWords:  glViewport mView setViewportSize IntSize Fennec GUIs init
%  LocalWords:  GLSurfaceView guardedRun checkThread CFA

### 2012-05-09 GUIErrorDetector/paper/related.tex
Mike's edits through section 4.1

Livshits et al.~\cite{Livshits:2005} presented a static analysis
to reason about reflective calls. The analysis
attempts to infer additional information stored in string constants to resolve
reflective calls statically. Their approach focuses on standard Java
reflection calls (e.g., \CodeIn{Class.forName}) instead of
framework-specific ones (e.g., \CodeIn{View.find-\\ViewById}).
TamiFlex~\cite{Bodden:2011}, a pure dynamic
approach, records all reflectively-created class instances
by intercepting JVM system calls, and re-inserts those recorded 
class into a program. However, TamiFlex requires a set of representative
program executions and is only sound with respect to the given executions.
Perhaps the closest work to our call graph
construction algorithm is Payet and Spoto's Julia
system~\cite{Payet:2011:SAA:2032266.2032299}. The Julia system
needs to first instrument Android's library code that performs the XML inflation,
and then replaces the \CodeIn{findViewById} call with the corresponding
object creation expressions. In addition, the Julia system does not handle
native methods when building call graphs. In contrast, our technique provides
annotation support for native methods, and our tool does not need a separate
pass of off-line instrumentation. The reflection-aware call graph is created
online by intercepting the standard call graph construction process.

@@

>>>>

Livshits et al.~\cite{Livshits:2005} presented a static analysis
to reason about reflective calls. The analysis
attempts to infer additional information stored in string constants to resolve
reflective calls statically. Their approach focuses on standard Java
reflection calls (e.g., \CodeIn{Class.forName}) instead of
framework-specific ones (e.g., \CodeIn{View.find\-ViewById}).
TamiFlex~\cite{Bodden:2011}, a pure dynamic
approach, records all reflectively-created class instances
by intercepting JVM system calls, and re-inserts those recorded 
class into a program. However, TamiFlex requires a set of representative
program executions and is only sound with respect to the given executions.
Perhaps the closest work to our call graph
construction algorithm is Payet and Spoto's Julia
system~\cite{Payet:2011:SAA:2032266.2032299}. The Julia system
needs to first instrument Android's library code that performs the XML inflation,
and then replaces the \CodeIn{findViewById} call with the corresponding
object creation expressions. In addition, the Julia system does not handle
native methods when building call graphs. In contrast, our technique provides
annotation support for native methods, and our tool does not need a separate
pass of off-line instrumentation. The reflection-aware call graph is created
online by intercepting the standard call graph construction process.

@@

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "guierror"
%%% TeX-command-default: "PDF"
%%% End:

%  LocalWords:  multithreaded Memon Xie Payet Spoto lockset Palsberg RTA
%  LocalWords:  Sundaresan Livshits forName ViewById TamiFlex Spoto's
%  LocalWords:  findViewById

### 2012-05-09 GUIErrorDetector/paper/technique.tex
revise more
Mike's edits through section 4.1

\vspace{1mm}
\noindent {\textsc{\textbf{Definition 1 (UI Thread).}}} {The UI thread
is a special thread created by the GUI framework during
GUI initialization. After the GUI becomes visible, the UI thread
takes charge of the application to handle events from the GUI.
It spawns new threads to process lengthy operations in the background. }
\vspace{1mm}

\noindent {\textsc{\textbf{Assumption 1.}}} {We assume that each multithreaded
GUI application has a single global UI thread. This is true for applications
built on top of GUI frameworks adopting the single-GUI-thread rule. The
only exception is that 
 an application may fork a new process to launch another
application with its own UI thread. In that case, we require the
launched application to be analyzed separately.}\vspace{1mm}

@@

\noindent {\textsc{\textbf{Assumption 2.}}} { We assume that each non-UI
thread is (transitively) spawned by the UI thread. Under this assumption,
we ignore all non-UI threads created by the GUI framework
before the UI thread has been initialized. That is, all post-initialization
GUI work occurs in the UI thread. Once the GUI is visible, the
application is driven by events, which are always handled in the UI thread.
We believe this assumption is reasonable, since if a non-UI thread 
spawned during pre-initialization GUI work accesses a GUI object, an exception becomes
immediately apparent and the whole application may abort even before the
GUI is visible. This is highly unlikely for fielded GUI
applications.
}\vspace{1mm}

@@

\noindent  {\textsc{\textbf{Definition 4 (Safe UI Method).}}} {GUI frameworks that
adopt the single-GUI-thread rule must provide methods to permit non-UI threads
to run code in the UI thread. We call such methods \textit{safe UI methods}, since
they can be invoked safely by any thread.}\vspace{1mm}

\noindent  {\textsc{\textbf{Definition 5 (Invalid Thread Access).}}} {An invalid-thread-access error occurs when the UI thread may spawn a non-UI thread, and there
exists a path from the non-UI thread's \CodeIn{start} method to any UI-accessing method
without going through any safe UI method. }\vspace{1mm}

@@

Figure~\ref{fig:detectalgorithm} shows the algorithm for detecting
potential invalid thread access errors. Our algorithm uses a
static call graph as the program representation for a multithreaded
GUI application. A Java call graph represents calling relationships
between methods. Specifically, each node represents a method and each
edge ($f$, $g$) indicates that method $f$ may call method $g$.
A theoretically ideal call graph can be defined as the union of the
dynamic call graphs over all possible executions of the program. 
A conservative, or sound, static call graph is a superset of
the ideal call graph; it over-approximates the
dynamic call graph of every possible execution. 
Based on call graph, our algorithm's reports are in terms of methods, as shown in Figure~\ref{fig:report}.

@@

Our algorithm in Figure~\ref{fig:detectalgorithm} %works as follows. It
first constructs a static call graph for the tested program (line 2),
then specifies entry nodes, UI-accessing nodes, and safe UI nodes
for it on lines 3, 4, and 5, respectively. Lines 3--5 are GUI framework-specific.
As an example, for a SWT desktop application, the entry nodes include
the single main method, UI-accessing nodes include methods
\CodeIn{Widget.checkWidget} and \CodeIn{Display.}\CodeIn{checkDevice},
and safe UI methods include two SWT helper methods \CodeIn{Display.asyncExec}  and
\CodeIn{Display.syncExec} for message passing. Section~\ref{sec:platforms}
explains the detailed instantiation for each supported framework.

@@

The algorithm in Figure~\ref{fig:detectalgorithm} uses BFS to search
for potential error-revealing paths, since BFS always returns the
shortest path to the UI-accessing node, permitting
smaller error reports. However,
other graph search strategies such as Depth-First Search (DFS) or
exhaustive path search can also be employed. In our experiment (Section~\ref{sec:search}),
we empirically compared three different graph search strategies, and demonstrated that
using BFS, the algorithm found more errors than DFS and
exhaustive path search.

@@

\begin{figure}[t]
\vspace{3mm}
\textbf{Input}: a Java program $\mathit{P}$\\
\textbf{Output}: a call graph $\mathit{cg}$\\
\vspace{-4mm}
\begin{algorithmic}[1]
\FOR{each $\mathit{expression}$ in $\mathit{P}$}
\IF{isReflectionCall($expression$)}
\STATE $\mathit{objectSet}$ $\leftarrow$ getAllObjectsThatMaybeCreated($\mathit{expression}$)
\STATE $\mathit{newExpr}$ $\leftarrow$ createObjectCreationExpression($\mathit{objectSet}$)
\STATE replace $\mathit{expression}$ with $\mathit{newExpr}$
\ENDIF
\ENDFOR
\STATE $\mathit{cg}$ $\leftarrow$ constructCallGraph($\mathit{P}$)
\RETURN $\mathit{cg}$
%\ENDWHILE
\end{algorithmic}
\vspace{-3mm}
\caption{A reflection-aware call graph construction algorithm. Lines
1--7 are a simple program transformation to replace reflection calls
with object creations, and line 8 builds the call graph using
an existing call graph construction algorithm. This algorithm
is parameterized by two methods on lines 2 and 3. 
How to instantiate it for the Android framework is presented in Section~\ref{sec:cg}.
} 
\label{fig:cgalgorithm}
\end{figure}

@@

This algorithm is parameterized by two methods on
lines 2 and 3. When instantiated for the Android framework,
the predicate isReflectionCall on line 2 returns true
if the $expression$ is a \CodeIn{findViewById(id)} method call, and the method
getAllObjectsThatMaybeCreated on line 3 parses the associated XML configuration
file to extract the class declaration
 corresponding to the given \CodeIn{id} value. If the \CodeIn{id}
value is dynamically generated, this method will conservatively
return instances of all subclasses of the declared type.
%On line 8, any existing call graph construction algorithm can be applied.

@@

A static analysis can check possible error paths without executing the code; but may report
paths that do not actually exist (false positives) or multiple paths
 that have the same error cause (redundant warnings). We devise
5 error filters below to remove likely false positives and redundant warnings.
The first 2 filters are sound in that they will not filter real bugs, and
the other 3 filters are based on heuristics. Orthogonally, filters 2
and 3 are for reducing false positives, and filters 1, 4, and 5
are for reducing redundant warnings.

@@

\textbf{1. Filter Lexically Redundant Reports}. One reported method call
chain can lexically subsume another one. For example, suppose that two
reported method call chains; \CodeIn{a()} $\rightarrow$ \CodeIn{b()}
$\rightarrow$ \CodeIn{c()} and 
\CodeIn{d()} $\rightarrow$ \CodeIn{a()} $\rightarrow$ \CodeIn{b()} $\rightarrow$ \CodeIn{c()},
both lead to a potential error, since $\CodeIn{d()}$ and \CodeIn{a()}
are two distinct entry methods. The second call chain can
be removed without reducing the error detection capability, since
the first chain reveals the same error and is shorter 
for programmers to interpret.

%\item
\textbf{2. Filter Reports with User-Annotated Methods}. Users are
permitted to explicitly annotate specific methods
that will never trigger an error.
%For example, the Android GUI framework provides a utility method 
%\CodeIn{runOnUIThread} to check whether the current thread is the UI thread
%before executing the code. If the current thread is not the UI thread,
%this method will be executed on the UI thread via message passing.
%Thus,
A reported method call chain containing an annotated  method
can be safely removed.

In our experiments, 
we produced a list of \filternum annotated methods for two
subjects (MyTracks and Fennec), since they employ a customized pattern
to interact with the GUI framework. For example, in MyTracks, all non-UI
threads are initialized via using the library method \CodeIn{android.os.}\discretionary{}{}{}\CodeIn{handler.handleCallback}.
This method can be invoked from any thread and will never cause an
invalid thread access error, because it checks whether the current
thread is the UI thread or not before execution. If not,
this method will be executed asynchronously via
message passing.

@@

\textbf{3. Filter Reports Containing Library Calls}. A
method call chain containing certain library calls
like \CodeIn{Runtime.}\CodeIn{shutDown} are unlikely to
be buggy. For example, the \CodeIn{Runtime.}\CodeIn{shutDown}
method is called when JVM terminates, and uses multithreading
to dispose all GUI objects when the program exits.
We encoded a list of 25 common library calls by proposing
a basic set of library calls that seemed %natural and
generally applicable, based on our programming experience.
We later added other calls we found helpful in analyzing
programs on different GUI frameworks and that we believed would be generally useful.
%We did this only between experiments rather than biasing experiments by tuning
%our tool to specific programs.

\textbf{4. Filter Report with the Same Head Methods from the Entry Node to \CodeIn{Thread.start()}}. A method can call
multiple methods that access GUI objects, such as:

\pagebreak[3]
\vspace{-2mm}                   %
\begin{CodeOut}
\begin{alltt}
     public void m() \ttlcb
         accessUIObject1();
         accessUIObject2();
     \ttrcb
\end{alltt}
\end{CodeOut}
\vspace{-2mm}
If method \CodeIn{m()} is invoked by a non-UI thread, method call chains that
only differ in the last few method nodes may be reported separately, such as:

@@

In fact, these two chains may have the same error root cause: knowing that
method \CodeIn{m()} is called by a non-UI thread is sufficient to understand
the error.
This filter compares two reported chains that have the same
head methods from the entry node to \CodeIn{Thread}.\CodeIn{start()},
and removes the longer one.
%$k$ ($k$ is user-settable with a default value 5 as used in our
%experiments) head nodes, and removes the longer one.

%\item
\textbf{5. Filter Report with the Same Tail Methods from \CodeIn{Thread}.
\CodeIn{start()}
to the UI-accessing Method}. A method
can have multiple callers, so method call chains with the same tail are likely
to just represent different ways to trigger the same error. This filter compares 
two reported method call chains that share the same 
tail from \CodeIn{Thread.start()} to the UI-accessing methods, and
removes the longer one. Using filters 4 and 5 together results in
only one error per \CodeIn{Thread.start()}.

@@

As we will show in our experiments (Section~\ref{sec:filters}), these filters
work remarkably well by reducing over 99\% of the reported warnings. In addition,
we also found that using sound filters alone was insufficient. The overwhelming
number of warnings after sound filters were applied motivated the use
of heuristic filters.

@@

>>>>

\vspace{1mm}
\noindent {\textsc{\textbf{Definition 1 (UI Thread).}}} {The UI thread
is a special thread created by the GUI framework during
GUI initialization. After the GUI becomes visible, the UI thread
takes charge of the application to handle events from the GUI\@.
It spawns new threads to process lengthy operations in the background. }
\vspace{1mm}

\noindent {\textsc{\textbf{Assumption 1.}}} {We assume that each multithreaded
GUI application has a single UI thread. This is true for applications
built on top of GUI frameworks adopting the single-GUI-thread rule. The
only exception is that 
 an application may fork a new process to launch another
application with its own UI thread. In that case, we require the
launched application to be analyzed separately.}\vspace{1mm}

@@

\noindent {\textsc{\textbf{Assumption 2.}}} { We assume that each non-UI
thread is (transitively) spawned by the UI thread. Under this assumption,
we ignore all non-UI threads created by the GUI framework
before the UI thread has been initialized. That is, we assume all post-initialization
GUI work occurs in the UI thread. Once the GUI is visible, the
application is driven by events, which are always handled in the UI thread.
We believe this assumption is reasonable, since if a non-UI thread 
spawned during pre-initialization GUI work accesses a GUI object, an exception becomes
immediately apparent and the whole application may abort even before the
GUI is visible. This is highly unlikely for fielded GUI
applications.
}\vspace{1mm}

@@

\noindent  {\textsc{\textbf{Definition 4 (Safe UI Method).}}} {GUI frameworks that
adopt the single-GUI-thread rule must provide methods to permit non-UI threads
to run code in the UI thread, typically by sending a message to the UI thread.
We call such methods \textit{safe UI methods}, since
they can be invoked safely by any thread.}\vspace{1mm}

\noindent  {\textsc{\textbf{Definition 5 (Invalid Thread Access).}}} {An invalid-thread-access error results when the UI thread may spawn a non-UI thread, and there
exists a path from the non-UI thread's \CodeIn{start} method to any UI-accessing method
without going through any safe UI method. }\vspace{1mm}

@@

Figure~\ref{fig:detectalgorithm} shows the algorithm for detecting
potential invalid thread access errors. Our algorithm uses a
static call graph as the program representation for a multithreaded
GUI application. A Java call graph represents calling relationships
between methods. Specifically, each node represents a method and each
edge ($f$, $g$) indicates that method $f$ may call method $g$.
A theoretically ideal call graph is the union of the
dynamic call graphs over all possible executions of the program. 
A conservative, or sound, static call graph is a superset of
the ideal call graph; it over-approximates the
dynamic call graph of every possible execution. 
Because it is 
based on the call graph, our algorithm's reports are in terms of methods, as shown in Figure~\ref{fig:report}.

@@

Our algorithm in Figure~\ref{fig:detectalgorithm} %works as follows. It
first constructs a static call graph for the tested program (line 2),
then specifies entry nodes, UI-accessing nodes, and safe UI nodes
for it on lines 3, 4, and 5, respectively. Lines 3--5 are GUI-framework-specific.
As an example, for a SWT desktop application, the entry nodes include
the single main method, UI-accessing nodes include methods
\CodeIn{Widget.checkWidget} and \CodeIn{Display.}\CodeIn{checkDevice},
and safe UI methods include two SWT helper methods \CodeIn{Display.asyncExec}  and
\CodeIn{Display.syncExec} for message passing. Section~\ref{sec:platforms}
explains the detailed instantiation for each supported framework.

@@

The algorithm in Figure~\ref{fig:detectalgorithm} uses BFS to search
for potential error-revealing paths, since BFS always returns the
shortest path to the UI-accessing node, permitting
smaller error reports. However,
other graph search strategies such as Depth-First Search (DFS) or
exhaustive path search can also be employed. In our experiment (Section~\ref{sec:search}),
we empirically compared three different graph search strategies, and demonstrated that
using BFS, the algorithm found more errors than DFS and was more practical than
exhaustive path search.

@@

\begin{figure}[t]
\vspace{3mm}
\textbf{Input}: a Java program $\mathit{P}$\\
\textbf{Output}: a call graph $\mathit{cg}$\\
\vspace{-4mm}
\begin{algorithmic}[1]
\FOR{each $\mathit{expression}$ in $\mathit{P}$}
\IF{isReflectionCall($expression$)}
\STATE $\mathit{objectSet}$ $\leftarrow$ getAllObjectsThatMayBeCreated($\mathit{expression}$)
\STATE $\mathit{newExpr}$ $\leftarrow$ createObjectCreationExpression($\mathit{objectSet}$)
\STATE replace $\mathit{expression}$ with $\mathit{newExpr}$
\ENDIF
\ENDFOR
\STATE $\mathit{cg}$ $\leftarrow$ constructCallGraph($\mathit{P}$)
\RETURN $\mathit{cg}$
%\ENDWHILE
\end{algorithmic}
\vspace{-3mm}
\caption{A reflection-aware call graph construction algorithm. Lines
1--7 are a simple program transformation to replace reflection calls
with object creations, and line 8 builds the call graph using
an existing call graph construction algorithm. This algorithm
is parameterized by two methods on lines 2 and 3. 
How to instantiate it for the Android framework is presented in Section~\ref{sec:cg}.
} 
\label{fig:cgalgorithm}
\end{figure}

@@

This algorithm is parameterized by two methods on
lines 2 and 3. When instantiated for the Android framework,
the predicate isReflectionCall on line 2 returns true
if the $expression$ is a \CodeIn{findViewById(id)} method call, and the method
getAllObjectsThatMayBeCreated on line 3 parses the associated XML configuration
file to extract the class declaration
 corresponding to the given \CodeIn{id} value. If the \CodeIn{id}
value is dynamically generated, this method will conservatively
return instances of all subclasses of the declared type.
%On line 8, any existing call graph construction algorithm can be applied.

@@

A static analysis can check possible error paths without executing the
code, but it may report
paths that do not actually exist (false positives) or multiple paths
that have the same error cause (redundant warnings). We devised
5 error filters to remove likely false positives and redundant warnings.
The first 2 filters are sound in that they will not filter real bugs, and
the other 3 filters are based on heuristics. Orthogonally, filters 2
and 3 are for reducing false positives, and filters 1, 4, and 5
are for reducing redundant warnings.

@@

\textbf{1. Filter Lexically Redundant Reports}. One reported method call
chain can lexically subsume another one. For example, suppose that two
reported method call chains, \CodeIn{a()} $\rightarrow$ \CodeIn{b()}
$\rightarrow$ \CodeIn{c()} and 
\CodeIn{d()} $\rightarrow$ \CodeIn{a()} $\rightarrow$ \CodeIn{b()} $\rightarrow$ \CodeIn{c()},
both lead to a potential error, since $\CodeIn{d()}$ and \CodeIn{a()}
are two distinct entry methods. The second call chain can
be removed without missing any errors, since
the first chain reveals the same error and is shorter 
for programmers to interpret.

%\item
\textbf{2. Filter Reports with User-Annotated Methods}. Users are
permitted to explicitly annotate specific methods
that will never trigger an error.
%For example, the Android GUI framework provides a utility method 
%\CodeIn{runOnUIThread} to check whether the current thread is the UI thread
%before executing the code. If the current thread is not the UI thread,
%this method will be executed on the UI thread via message passing.
%Thus,
Assuming the user annotations are accurate,
a reported method call chain containing an annotated method
can be safely removed.

In our experiments, 
we produced a list of \filternum annotated methods for two
subjects (MyTracks and Fennec), since they employ a customized pattern
to interact with the GUI framework. For example, in MyTracks, all non-UI
threads are initialized via using the library method \CodeIn{android.os\-.handler.handleCallback}.
This method can be invoked from any thread and will never cause an
invalid thread access error, because it checks whether the current
thread is the UI thread or not before execution. If not,
this method will use safe UI methods to run the code in the UI thread.

@@

\textbf{3. Filter Reports Containing Library Calls}. A
method call chain containing certain library calls
like \CodeIn{Runtime.}\CodeIn{shutDown} are unlikely to
be buggy. For example, the \CodeIn{Runtime.}\CodeIn{shutDown}
method is called when the JVM terminates, and uses multithreading
to dispose all GUI objects when the program exits.
% We encoded a list of 25 common library calls by proposing
% a basic set of library calls that seemed %natural and
% generally applicable, based on our programming experience.
% We later added other calls we found helpful in analyzing
% programs on different GUI frameworks and that we believed would be generally useful.
Our experiments use a list of 25 such library calls.
%We did this only between experiments rather than biasing experiments by tuning
%our tool to specific programs.

\textbf{4. Filter Reports with the Same Head Methods from the Entry Node to \CodeIn{Thread.start()}}. A method can call
multiple methods that access GUI objects, such as:

\pagebreak[3]
\vspace{-2mm}                   %
\begin{CodeOut}
\begin{alltt}
     public void m() \ttlcb
         accessUIObject1();
         accessUIObject2();
     \ttrcb
\end{alltt}
\end{CodeOut}
\vspace{-2mm}
If method \CodeIn{m()} is invoked by a non-UI thread, our algorithm will
report method call chains that
only differ in the last few method nodes, such as:

@@

\noindent
These two chains may have the same error root cause: knowing that
method \CodeIn{m()} is called by a non-UI thread is sufficient to understand
the error.
This filter compares two reported chains that have the same
head methods from the entry node to \CodeIn{Thread}.\CodeIn{start()},
and removes the longer one.
%$k$ ($k$ is user-settable with a default value 5 as used in our
%experiments) head nodes, and removes the longer one.

%\item
\textbf{5. Filter Reports with the Same Tail Methods from \CodeIn{Thread}.
\CodeIn{start()}
to the UI-accessing Method}. A method
can have multiple callers, so method call chains with the same tail are likely
to just represent different ways to trigger the same error. This filter compares 
two reported method call chains that share the same 
tail from \CodeIn{Thread.start()} to the UI-accessing methods, and
removes the longer one. Using filters 4 and 5 together results in
only one error per \CodeIn{Thread.start()}.

@@

In our experiments (Section~\ref{sec:filters}), these filters remove
99.96\% of the reported warnings.
We also found that using sound filters alone was insufficient, because an overwhelming
number of warnings remained.  This motivated our use
of heuristic filters.

@@

%  LocalWords:  multithreaded pre SWT checkWidget checkDevice asyncExec
%  LocalWords:  syncExec BFS createErrorReport LayerView init ViewRoot
%  LocalWords:  checkThread constructCallGraph getEntryNodes newError
%  LocalWords:  getUIAccessingNodes getSafeUINodes getReachableStarts
%  LocalWords:  uiAccessingNodes safeUINodes worklist objectSet newExpr
